---
title: "A4"
author: "Adrien Dinzey, 7842736"
date: "23/04/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Q1

load in data:
```{r}
library('readxl')
data1<-read_excel("A4-datasets.xlsx")
```

Using the cannonical link (that is the recirpocal for Gamma dsitribution)
```{r}
data1.model<-glm(y~x1+x2+x3,family=Gamma(),data=data1)
summary(data1.model)
yhat<-data1.model$fit
dev_res<-residuals(data1.model,c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data1$x1,dev_res)
plot(data1$x2,dev_res)
plot(data1$x3,dev_res)

```

Using the log link
```{r}
data1.model<-glm(y~x1+x2+x3,family=Gamma(link=log),data=data1)
summary(data1.model)
yhat<-data1.model$fit
dev_res<-residuals(data1.model,c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data1$x1,dev_res)
plot(data1$x2,dev_res)
plot(data1$x3,dev_res)

```

We see that the residuals seem to be distributed the same way using either link. The QQ plot of the canonical link seems to look like a normal qq plot while the log link qq plot has a bit of a bend near the middle and is pushed upwards.

# Q2
```{r}
data2<-read_excel("A4-datasets.xlsx",sheet = 2)
```

```{r}
data2.model<-glm(y~x,family = "poisson",data2)
summary(data2.model)
yhat <- data2.model$fit
dev_res <- residuals(data2.model, c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data2$x,dev_res)
```
We see that the residuals plotted against yhat seems to look like some sort of F distribution while when plotted against x it looks like an upside down parabola. Its possible that a power link would be better suited for this.

# Q3
```{r}
data3<-read_excel("A4-datasets.xlsx",sheet=3)
```

```{r}
data3.model<-glm(y~x1+x2+x3,family=Gamma(link=power(1/2)),data=data3)
summary(data3.model)
yhat<-data3.model$fit
dev_res<-residuals(data3.model,c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data3$x1,dev_res)
plot(data3$x2,dev_res)
plot(data3$x3,dev_res)
```

# Q4

```{r}
data4<-read_excel("A4-datasets.xlsx",sheet = 4)
```

Using canonical link
```{r}
data4.model<-glm(NO~x1+x2+x3,family = poisson,data4)
summary(data4.model)
yhat <- data4.model$fit
dev_res <- residuals(data4.model, c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data4$x1,dev_res)
plot(data4$x2,dev_res)
plot(data4$x3,dev_res)
```
Using log link
```{r}
data4.model<-glm(NO~x1+x2+x3,family = poisson(link=log),data4)
summary(data4.model)
yhat <- data4.model$fit
dev_res <- residuals(data4.model, c="deviance")
qqnorm(dev_res)
plot(yhat,dev_res)
plot(data4$x1,dev_res)
plot(data4$x2,dev_res)
plot(data4$x3,dev_res)
```

We see we get the exact same answer using log link or canonical link here because the canonical link for poisson distribution is the log link.
